\chapter{Introduction}
\label{chap:Chapter1}

This chapter provides context for automating alert triage and its significance in enhancing security operations in a SOC, particularly for managing high volumes of security alerts. It begins with defining the problem and detailing the challenges associated with manual alert analysis and the critical requirements for automation. First, the objectives of this thesis are presented; second, this project's expected outcomes, including efficiency and accuracy improvements, are outlined, along with the selected approach for integrating an existing ML model to automate alert classification. Finally, an overview of the report's structure is provided.

\section{Context}

In today's world, cybersecurity has become essential for organizations of all sizes. \parencite{Vielberth2020} The rapid advancement of technology, particularly artificial intelligence, has fostered innovation and growth and equipped cybercriminals with increasingly sophisticated tools and techniques. Cyber threats have surged over recent years, and this trend is expected to continue, with global cybersecurity spending projected to rise from \$6 trillion to \$30 trillion by 2030. \parencite{SOVILJ2020113577}

As these threats become more complex and frequent, \gls{SOC} have emerged as the primary line of defense. \gls{SOC}s are crucial in monitoring, detecting, and responding to cyber threats, ensuring the protection of organizations' core digital assets—data, applications, and infrastructure. By combining human expertise with advanced technology, \gls{SOC}s proactively defend systems, respond to incidents, and work to minimize the impact of potential breaches, safeguarding essential operations. \parencite{Jalalvand2024}

However, managing the overwhelming number of alerts generated daily within \gls{SOC}s presents a significant challenge. Analysts must review and respond to thousands of tickets, a process that can quickly lead to burnout, as stated by \textcite{Tines2023}, especially when the majority of alerts are false positives. A study found that many organizations receive over 10,000 alerts daily, with more than 50\% being false positives. \parencite{CriticalStart2019}

One of the stages in the \gls{SOC} workflow with a higher alert volume is the initial triage process. At this stage, Tier 1 analysts are responsible for gathering raw data, assessing alarms, and determining the urgency and nature of each alert. For every ticket, they must evaluate whether the alert is valid or a false positive, assign priority based on the alert’s severity, and identify any potential high-risk incidents. \parencite{Vielberth2020} This process demands significant time and effort, placing a heavy mental load on analysts as they deal with large volumes of information.

Integrating automation in the triage process, particularly for determining alert criticality and classification, could substantially reduce analyst fatigue. \gls{ML} techniques are up-and-coming for processing large data volumes. They facilitate the classification and prioritization of alerts, enabling analysts to focus on high-priority threats. \parencite{Jalalvand2024}

This work explores how \gls{SOC}s can use automation and \gls{ML} to streamline alert triage, improve \gls{SOC} efficiency, and more effectively protect organizations from an ever-evolving cyber threat landscape.

%-------------------------------------------------------------------------------
\section{Problem}

The main objective of this work is to develop an independent program that automates the triage of security alerts within a Security Operations Center (SOC) environment. This automation aims to enhance the overall efficiency and accuracy of the triage process, reduce the number of false positives, and alleviate the workload during the initial stages of the SOC workflow.

Improving the speed of analyzing initial alerts is crucial in a Security Operations Center (SOC) workflow. Faster analysis enables SOC analysts to respond to incidents quickly, which helps minimize the time that threats remain undetected. Without automation, this initial alert analysis depends heavily on manual inspection. Analysts face the challenge of sifting through a large volume of alerts to identify potential threats, which is time-consuming and susceptible to human error. 

This reactive, manual method increases response times and raises the risk of missing critical threats. Analysts must manually classify alerts, assess their severity, and determine appropriate response actions while managing a continuous influx of new alerts. This approach can strain resources and often leads to alert fatigue, making it difficult for analysts to prioritize and respond effectively.

Key metrics such as Mean Time to Detect (MTTD) and Mean Time to Respond (MTTR) can significantly improve by reducing the time required for initial analysis. These metrics are essential for maintaining strong cybersecurity defenses. Automation in alert triage allows for faster and more accurate threat identification, enabling analysts to focus on high-priority threats. This not only increases overall productivity but also reduces fatigue among analysts, while boosting the effectiveness of cybersecurity.

This efficiency results in a more manageable workload, decreased burnout, and greater job satisfaction for employees. For the organization, streamlined response times enhance the security posture, lower the risk of breaches, and ultimately protect critical assets more effectively.

%-------------------------------------------------------------------------------
\section{Objectives}

The objectives for this project are as follows:

\begin{enumerate}
    \item Literature search and review
    \begin{itemize}
        \item A research into the latest advancements in security analysis and a thorough analysis of the tools, methodologies, and approaches researchers use in SOC automation and machine learning applications.
    \end{itemize}
    
    \item Study of existing Machine Learning (ML) models
    \begin{itemize}
        \item An analysis of different models for alert triage in SOC environments and related fields, aiming to identify and improve the adopted approach.
    \end{itemize}
    
    \item Collection of security alert datasets
    \begin{itemize}
        \item A compilation of accessible datasets on security alerts organized to supply training data for the selected ML model.
    \end{itemize}

    \item Integration with SIEM and other data sources
    \begin{itemize}
        \item Connect the SIEM and relevant data sources for real-time data ingestion and analysis.
    \end{itemize}
    
    \item Developing a custom dashboard for analysts to submit feedback on the machine learning model.
    \begin{itemize}
        \item While the ML model will be trained with existing datasets before deployment, it is still beneficial to provide a method for analysts to continue training the model over time.
    \end{itemize}
    
    \item Testing and evaluation of the selected ML model
    \begin{itemize}
        \item The selected ML model will be tested with live data, followed by an analysis to assess its effectiveness in categorizing alerts and minimizing false positives.
    \end{itemize}
\end{enumerate}


\section{Research Contextualization}



%-------------------------------------------------------------------------------
\section{Dissertation Structure}

\textit{
    [ 
        Structure
    ]
}

%-------------------------------------------------------------------------------
%---------
%